---
title: "Analysis 2.0"
author: "AMH"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Amy's packages and working directory
```{r}
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(rcompanion)

setwd("C:/Users/hrusk/Dropbox (Smithsonian)/2021_SummerInterns/AM")

dat <- read.csv("Heat_herbivory_alldata_v2.csv")


```


Annika just messin' around 
```{r}
shawty <- 3*7*16
print(shawty)
```

Annika's packages and working directory
```{r}

```

Data wrangling/cleaning
```{r}
##remove spaces from dataset#
dat2 <- dat %>%
  mutate(across(where(is.factor), str_remove_all, pattern = fixed (" ")))

#####################################################
#################### height data ####################
#####################################################

#create dataset for height graphing and analysis#
height <- dat2 %>%
  select(Date, Species, Temperature, Plant_number, Height_cm)
height$Height_cm <- as.numeric(height$Height_cm)
height2 <- na.omit(height)

#test height_cm is numeric and set to numeric#
is.numeric(height2$Height_cm)
height2$Height_cm <- as.numeric((height2$Height_cm))

#height last day#
height3 <- height2 %>%
  filter(Date == "9-Aug")
height3 <- na.omit(height3)
height3$Height_cm <- as.numeric(height3$Height_cm)

#height first day#
height4 <- height %>%
  filter(Date == "8-Jul")
height4$Height_cm <- as.numeric((height4$Height_cm))

#merge height of last and first date and create a new column for change in height#
height5 <- merge(height3, height4, by = "Plant_number")
height6 <- na.omit(height5)
height6$Height_cm.x <- as.numeric((height6$Height_cm.x))
height6$Height_cm.y <- as.numeric((height6$Height_cm.y))

height7 <- height6 %>% 
  mutate(delta= height6$Height_cm.x - height6$Height_cm.y)

######################################################
#################### biomass data ####################
######################################################

biomass <- dat2 %>%
  select(Date, Species, Temperature, Plant_number, abg_biomass_g, blg_biomass_g)


biomass2 <- biomass
biomass2[biomass2 == "" | biomass2 == "N/A"] <- NA
biomass2$blg_biomass_g <- as.numeric(biomass2$blg_biomass_g)

abg_biomass <- biomass2 %>%
  filter(abg_biomass_g > 0)

blg_biomass <- biomass2 %>%
  filter(blg_biomass_g > 0)

total <- blg_biomass %>%
  mutate(total_biomass_g = blg_biomass$abg_biomass_g + blg_biomass$blg_biomass_g)
  
  
```

BarStats Function
```{r}

barGraphStats <- function(data, variable, byFactorNames) {
  count <- length(byFactorNames)
  N <- aggregate(data[[variable]], data[byFactorNames], FUN=length)
  names(N)[1:count] <- byFactorNames
  names(N) <- sub("^x$", "N", names(N))
  mean <- aggregate(data[[variable]], data[byFactorNames], FUN=mean)
  names(mean)[1:count] <- byFactorNames
  names(mean) <- sub("^x$", "mean", names(mean))
  sd <- aggregate(data[[variable]], data[byFactorNames], FUN=sd)
  names(sd)[1:count] <- byFactorNames
  names(sd) <- sub("^x$", "sd", names(sd))
  preSummaryStats <- merge(N, mean, by=byFactorNames)
  finalSummaryStats <- merge(preSummaryStats, sd, by=byFactorNames)
  finalSummaryStats$se <- finalSummaryStats$sd / sqrt(finalSummaryStats$N)
  return(finalSummaryStats)
  
}
```

Plant Height Graphs and Analyses
Plant height over all weeks
```{r}
##Creating colourblind palette###
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

##saved graph as variable to edit theme below##
g1 <- ggplot(data = barGraphStats(data = height2, variable = "Height_cm", byFactorNames = c("Species", "Temperature")), aes(x=Species, y=mean, fill = Temperature)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
    xlab("Plant Species") + ylab("Plant Height (cm)") +
    scale_fill_manual(values = cbPalette) 

##setting theme to remove background grid and add axis lines##    
g1 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 16),axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
           axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

ggsave("plant_height_allweeks.png")

```


Height on final day graph
```{r}
##saved graph as variable to edit theme below##
g2 <- ggplot(data = barGraphStats(data = height3, variable = "Height_cm", byFactorNames = c("Species", "Temperature")), aes(x=Species, y=mean, fill = Temperature)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
    xlab("Plant Species") + ylab("Plant Height (cm)") +
    scale_fill_manual(values = cbPalette) 

##setting theme to remove background grid and add axis lines##    
g2 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 16),axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)), axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

ggsave("height_finalday.png")
```
Delta height graph
```{r}
g2 <- ggplot(data = barGraphStats(data = height7, variable = "delta", byFactorNames = c("Species.x", "Temperature.x")), aes(x=Species.x, y=mean, fill = Temperature.x)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
    xlab("Plant Species") + ylab("Change in Plant Height (cm)") +
    scale_fill_manual(values = cbPalette) 

##setting theme to remove background grid and add axis lines##    
g2 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 16),axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)), axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

ggsave("height_delta.png")
```
Stats on plant height
```{r}

#test for normality, definitly not normal#
shapiro.test(height2$Height_cm)
ggqqplot(height2$Height_cm)

#logtransform and checking normality of height
ggqqplot(log(height2$Height_cm))

##packages for glm, glmms#
library(lme4)
library(nlme)
library(lattice)
library(effects)

mod1 <- lme(Height_cm ~ Temperature*Species, random = ~1|Date, data = height2)
anova(mod1)

mod2 <- glm(Height_cm ~ Temperature*Species, data = height3)
anova(mod2)

mod3 <- glm(delta ~ Temperature.x*Species.x, data = height7)
anova.glm(mod3)

```

Biomass graphs and analyses
Total biomass graph
```{r}
g3 <- ggplot(data = barGraphStats(data = total, variable = "total_biomass_g", byFactorNames = c("Species", "Temperature")), aes(x=Species, y=mean, fill = Temperature)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
    xlab("Plant Species") + ylab("Total Biomass (g)") +
    scale_fill_manual(values = cbPalette) 

##setting theme to remove background grid and add axis lines##    
g3 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 16),axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)), axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))


```

Aboveground Biomass graph
```{r}
g4 <- ggplot(data = barGraphStats(data = abg_biomass, variable = "abg_biomass_g", byFactorNames = c("Species", "Temperature")), aes(x=Species, y=mean, fill = Temperature)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
    xlab("Plant Species") + ylab("Aboveground Biomass (g)") +
    scale_fill_manual(values = cbPalette) 

##setting theme to remove background grid and add axis lines##    
g4 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 16),axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)), axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

```
Belowground Biomass
```{r}
g5 <- ggplot(data = barGraphStats(data = total, variable = "blg_biomass_g", byFactorNames = c("Species", "Temperature")), aes(x=Species, y=mean, fill = Temperature)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
    xlab("Plant Species") + ylab("Belowground Biomass (g)") +
    scale_fill_manual(values = cbPalette) 

##setting theme to remove background grid and add axis lines##    
g5 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 16),axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)), axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

```

